---
output:
  word_document:
    reference_docx: Style_reference.docx
  pdf_document: default
params:
  lake: WTL
  year: '2018'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)


#Librarys
#Functions
---

# Libarys for data manipulation, import functions
library(tidyverse) # This loads dplyr which is primarily needed here.
# dplyr also helps with calculation. 
# Very easy to understand the syntax in my opinion
# Optionally instead of loading tidyverse, you can load dplyr
# library(dplyr)
library(lubridate)



# Import data, mastersheet excel file
zimported <- read.csv("MasterSheet.csv")

# NA are -999 so convert to NULL or NA
zimported[zimported==-999] <- NA 

# Drop id collumn, convert dates to something R understands
# Calculate total nitrogen and tn:tp ratio 
Master_Sheet <- zimported %>% 
  select(-id) %>% # Dropping ID collumn 
  mutate(Dates = ymd(zimported$Date)) %>% 
  mutate(TN = NO3 + TKN) %>%
  mutate(TNTP = TN/TP)

#######################

# Import precipitation data and lake stations for each lake
# Imported from QGIS output
zghcnd <- read.csv("GHCND.csv")
zzghcnd <- read.csv("zzghcnd.csv")

# Join precip data to lake by station
zjoined <- zghcnd %>% 
  full_join(zzghcnd, by = c("STATION" = "station")) %>% 
  drop_na(lk_code)

# Filter out collumns and add dates and months
zNOAA <- zjoined %>% 
  select(date = DATE, prcp = PRCP, LK_CODE = lk_code, tobs = TOBS) %>%
  mutate(Dates = ymd(date)) %>%
  select(Dates, prcp, LK_CODE, tobs) %>%
  mutate(Months.rain = (month(Dates, label=TRUE, abbr = FALSE)))


#######################


# Ecoregion Import
ecoregion <- read.csv(file="ecoregion.csv")

# Lake information. Takes lat and long, county and land use from qgis/arcgis calculation
Lake_info_comprehensive <- read.csv(file="Lake_info_comprehensive.csv")

# Imports hobo data, they are averaged in excel already.
# Averaged each month for every lake. 
zhobo <- read.csv(file="Hobo_MONTHLY.csv")

# Calculate land use percentages from Lake info sheet since
# they are in acres
info <- Lake_info_comprehensive %>% 
  rowwise() %>% 
  mutate(sum = sum(Water,Developed,Barren,Forest,Shrubs,
                   Herbaceous,Agriculture,Wetlands)) %>% 
  mutate_at(vars(Water:Wetlands), funs(./sum)) 

# Data Manipulation

# Order months by month instead of alphabetical. 
# Also calculate total Sum from SPATTS 
# Added a time interval for temporal joins

MonthOrdered <- as.data.frame(Master_Sheet) %>% 
  mutate(month = factor(month.name[Month], levels = month.name)) %>% 
  arrange(month) %>% 
  mutate_at(vars(
    S_D_Asp3_RR, S_MC_RR, S_Nodul,
    S_MC_YR, S_MC_HtyR, S_MC_LR,
    S_D_Asp3_RR, S_MC_HilR, S_MC_WR,
    S_MC_LA, S_MC_LY, S_MC_LW,
    S_MC_LF), funs(.*45/3)) %>% 
  # Converted spatts into nanogram per gram of resin
  rowwise() %>% 
  mutate(S_SUM = 
           sum(S_D_Asp3_RR, S_MC_RR, S_Nodul, 
               S_MC_YR, S_MC_HtyR, S_MC_LR,
               S_D_Asp3_RR, S_MC_HilR, S_MC_WR,
               S_MC_LA, S_MC_LY, S_MC_LW,
               S_MC_LF)) %>%
  mutate(fiveday = ymd(Dates - ddays(5)),
         threeday = ymd(Dates - ddays(3)),
         sevenday = ymd(Dates - ddays(7)) 
         ,thirtyday = ymd(Dates - ddays(30))) 

MonthOrdered[MonthOrdered==Inf] <- 0





#Precipitation Data Join
zRain <- full_join(MonthOrdered, 
                   zNOAA, 
                   by = c("LK_CODE" = "LK_CODE"))

#Filter rain data to match date ranges.
zNOAA3 <- zRain %>% rowwise() %>%
  filter(Dates.y >= as.Date(threeday) & 
           Dates.y <= as.Date(Dates.x))
zNOAA5 <- zRain %>% rowwise() %>% 
  filter(Dates.y >= as.Date(fiveday) &
           Dates.y <= as.Date(Dates.x))
zNOAA7 <- zRain %>% rowwise() %>% 
  filter(Dates.y >= as.Date(sevenday) &
           Dates.y <= as.Date(Dates.x))
zNOAA30 <- zRain %>% rowwise() %>% 
  filter(Dates.y >= as.Date(thirtyday) &
           Dates.y <= as.Date(Dates.x))

#Average by Day of sampling varied by lagged temprature dates. 
zzNOAA3 <- zNOAA3 %>% 
  group_by(LK_CODE, Mo) %>% 
  summarise(precip3 = mean(prcp, na.rm = TRUE),
            temp3 = mean(tobs, na.rm =TRUE))

zzNOAA5 <- zNOAA5 %>% 
  group_by(LK_CODE, Mo) %>% 
  summarise(precip5 = mean(prcp, na.rm = TRUE), 
            temp5 = mean(tobs, na.rm =TRUE))

zzNOAA7 <- zNOAA7 %>%
  group_by(LK_CODE, Mo) %>% 
  summarise(precip7 = mean(prcp, na.rm = TRUE), 
            temp7 = mean(tobs, na.rm =TRUE))

zzNOAA30 <- zNOAA30 %>% 
  group_by(LK_CODE, Mo) %>% 
  summarise(precip30 = mean(prcp, na.rm = TRUE),
            temp30 = mean(tobs, na.rm =TRUE))

# Surveyor information

zzsurveyor <- read.csv(file="surveyor.csv") 

#Ultimate Join of all tables
zUltimateSheet <- MonthOrdered %>%
  full_join(info,
            by = c("LK_CODE" = "LK_CODE")) %>%
  full_join(ecoregion, 
            by = c("LK_CODE" = "LK_CODE")) %>%
  full_join(zzNOAA3,
            by = c("LK_CODE" = "LK_CODE", "Mo" = "Mo")) %>%
  full_join(zzNOAA5,
            by = c("LK_CODE" = "LK_CODE", "Mo" = "Mo")) %>%
  full_join(zzNOAA7,
            by = c("LK_CODE" = "LK_CODE", "Mo" = "Mo")) %>%
  full_join(zzNOAA30, 
            by = c("LK_CODE" = "LK_CODE", "Mo" = "Mo")) %>%
  full_join(zhobo, 
            by = c("LK_CODE" = "LK_CODE", "Month" = "Month")) %>%
  full_join(zzsurveyor, 
            by = c("LK_CODE" = "LK_CODE")) %>%
  rename(hobotemp = TEMP, 
         hobolight = LIGH, 
         ecoregion = US_L3NAME) %>%
  mutate(LK_CODE = factor(LK_CODE),
         year = as.factor(year),
         network = factor(Network, levels = c("1", "2", "3", "4"),
                          exclude = TRUE),
         open = as.factor(Open_clos)) %>%
  rowwise() %>%
  # Omit all great lake sites. Lake Superior
  filter_all(all_vars(!grepl('SUP',.))) %>% 
  # Omit Lake Erie
  filter_all(all_vars(!grepl('ERI',.))) %>% 
  # Omit Lake St. Clair
  filter_all(all_vars(!grepl('STC',.))) %>% 
  # Omit Cullen Park 
  filter_all(all_vars(!grepl('CUL',.))) %>% 
  select(-county, -REACHCODE, -sum) 



# Convert all characters as factor
UltimateFactor <- as.data.frame(zUltimateSheet) %>% 
  mutate_if(is.character, as.factor) %>%
# Surveyors such as Ryan, Jason, Ben and Szlag team
# are converted to dummy variables as factor levels
  mutate(surveyor = factor(surveyor, levels = c(1,2,3,4))) 


# Modeling Log Transformed dataset. 
# Each variable is added by reportable detection limit
#Log transformed all data
LOGTransformed <- UltimateFactor %>% 
  mutate_at(vars(OP),
            funs(log10(.+0.003))) %>%
  mutate_at(vars(NO3), 
            funs(log10(.+0.04))) %>%
  mutate_at(vars(NH3),
            funs(log10(. + 0.006))) %>%
  mutate_at(vars(TP), 
            funs(log10(. + 0.002))) %>%
  mutate_at(vars(TKN),
            funs(log10(. + 0.07))) %>%
  mutate_at(vars(TN), 
            funs(log10(. + 0.116))) %>%
  mutate_at(vars(DO),
            funs(log10(. + 0.01))) %>%
  mutate_at(vars(turb, conduc),
            funs(log10(. + 0.01))) %>%
  mutate_at(vars(ELISA), 
            funs(log10(. + 0.15))) %>%
  mutate_at(vars(LkArea), 
            funs(log10(. + 1))) %>%
  mutate_at(vars(WtShArea), 
            funs(log10(. + 1))) %>%
  mutate_at(vars(phyco,chloro), 
            funs(log10(. + 0.01))) %>%
  mutate_at(vars(X16SRNA, CYANA, MCYE, CyrA, SxtA), 
            funs(log10(. + 20))) %>% 
  # qpcr results log transformed
  mutate_at(vars(Nodul:SUM), 
            funs(log10(. + 0.03))) %>% 
  # Mass Spec
  mutate_at(vars(Anatoxin:S_ELISA), 
            funs(log10(. + 1))) %>% 
  # SPATTS
  mutate_at(vars(precip3, precip5, precip7, 
                 precip30, hobolight), 
            funs(log10(.+1))) %>%
  # Drop variables not to be included in modeling
  select(-CYANA,-SxtA,-Sensored,-PPIA,-Sensored_1,
         -S_EL_CENS,-Dates, -month,
         -fiveday:-Network,-Open_clos,
         -Zebra,-network,
         -open, -Anatoxin, -Cylindro,-CyrA)


#Averaged by each lake, then log transformed
MCSUMresponseAverage <- UltimateFactor  %>%
  select(-CYANA,-SxtA,-Sensored,-PPIA,-Sensored_1,
         -S_EL_CENS,-Dates, -month,
         -fiveday:-Network,-Open_clos,
         -Zebra,-network,
         -open, -Anatoxin, -Cylindro,-CyrA) %>%
  group_by(LK_CODE) %>%
  summarise_all(funs(mean(., na.rm=TRUE))) %>%
  mutate_at(vars(OP), funs(log10(.+0.003))) %>%
  mutate_at(vars(NO3), funs(log10(.+0.04))) %>%
  mutate_at(vars(NH3), funs(log10(. + 0.006))) %>%
  mutate_at(vars(TP), funs(log10(. + 0.002))) %>%
  mutate_at(vars(TKN), funs(log10(. + 0.07))) %>%
  mutate_at(vars(TN), funs(log10(. + 0.116))) %>%
  mutate_at(vars(DO), funs(log10(. + 0.01))) %>%
  mutate_at(vars(turb, conduc), funs(log10(. + 0.01))) %>%
  mutate_at(vars(ELISA), funs(log10(. + 0.15))) %>%
  mutate_at(vars(LkArea), funs(log10(. + 1))) %>%
  mutate_at(vars(WtShArea), funs(log10(. + 1))) %>%
  mutate_at(vars(phyco,chloro), funs(log10(. + 0.01))) %>%
  mutate_at(vars(X16SRNA, MCYE), funs(log10(. + 20))) %>%
  mutate_at(vars(Nodul:SUM), funs(log10(. + 0.03))) %>% 
  #Mass Spec
  mutate_at(vars(S_D_Asp3_RR:S_SUM), funs(log10(. + 1))) %>% 
  # SPATTS
  mutate_at(vars(precip3, precip5, precip7,
                 precip30, hobolight), funs(log10(.+1))) %>%
 select(-LK_CODE:-doy, -Nodul:-MC_LF, 
         -ecoregion, -surveyor, -ELISA,
         -S_D_Asp3_RR:-S_ELISA)

########################################################################
########################################################################
########################################################################



```{r DataManipulation, include=FALSE}
## Introduction

  

## Microcystin and Aquatic Toxin Summary

Enzyme Linked Immunosorbent  Assay (ELISA) is a most commonly used analytical method for microcystin (MC) detection. Further data will demonstrate the validity of the test across multiple aquatic environments. We will be ultimately comparing ELISA with Mass Spectrometry (MS) results because it is the most precise and accurate method of quantification.  Different MC congeners were measured and added as a sum (Sum of MC Congeners) to be compared with ELISA.

The information collected reflects the conditions at the time of sampling, but toxins recovered from the SPATT bags are indicative of toxin levels during the weeks between sampling. The SPATT bags were deployed for the month-long period between sampling events and then collected. SPATT data are represented by the total toxins collected during the period. The reported concentrations are from the analytical procedures used on the SPATT bags and are not directly indicative of average toxin levels. There is no USEPA guidance for this procedure yet and we will use the 2017 and 2018 data to provide interim guidance.


## Grab Sample


```{r, fig.width=9}
 MonthOrdered %>% select(LK_CODE, Month, year, ELISA, SUM) %>% filter(LK_CODE == params$lake) %>%   gather(Analysis, toxin, ELISA, SUM) %>%
  ggplot(aes(x=Month)) + 
  geom_line(aes(y=toxin, group=Analysis, color=Analysis), linetype='solid') + 
  geom_point(aes(x=Month, y=toxin)) + 
  ylab("MC in ppb") +
  scale_x_continuous(breaks = c(7,8,9,10) ,label=c("July", "August", "September", "October"))+
  geom_hline(aes(yintercept = 4), color="#BB0000", linetype="dashed", show.legend = FALSE) + 
  scale_color_discrete(labels = c("ELISA", "Sum of \nMC Congeners")) +
  geom_text(aes(9,4.4, label = "EPA Guideline"), color="#BB0000", show.legend=FALSE) +
  labs(title = "Total MC from Grab Samples") +
  theme_classic() 
```




\newpage


## SPATTS




```{r, fig.width=9}
 MonthOrdered %>% select(LK_CODE, Month, year, S_D_Asp3_RR:S_MC_LF) %>% filter(LK_CODE == params$lake) %>%   gather(Congeners, toxin, S_D_Asp3_RR:S_MC_LF) %>%
  ggplot(aes(x=Month, group=Congeners)) + 
  geom_line(aes(y=toxin, linetype=Congeners, color=Congeners)) + 
  geom_point(aes(x=Month, y=toxin), show.legend = FALSE) + 

  ylab("Recovered ng of MC / g of HP20 resin") +
  scale_x_continuous(breaks = c(8,9,10) ,label=c( "August", "September", "October"), limit = c(8,10)) +
  scale_color_discrete(labels = c("D Asp3 RR", "D Asp3 LR", "MC HiLR", "MC HtyR", "MC LA", "MC LF", "MC LR", "MC LW", "MC LY", "MC RR", "MC WR", "MC YR", "Nodularin")) +
  scale_linetype_discrete(labels = c("D Asp3 RR", "D Asp3 LR", "MC HiLR", "MC HtyR", "MC LA", "MC LF", "MC LR", "MC LW", "MC LY", "MC RR", "MC WR", "MC YR", "Nodularin")) +
  labs(title = "MC Congeners from SPATTS") +
  theme_classic()

```







```{r, fig.width=9}
 MonthOrdered %>% select(LK_CODE, Month, year, S_ELISA, S_SUM) %>% filter(LK_CODE == params$lake) %>%   gather(Analysis, toxin, S_ELISA, S_SUM) %>%
  ggplot(aes(x=Month)) + 
  geom_line(aes(y=toxin, group=Analysis, color=Analysis), linetype='solid') + 
  geom_point(aes(x=Month, y=toxin)) + 
  ylab("Recovered ng of MC / g of HP20 resin") +
  scale_x_continuous(limits = c(8,10), breaks = c(8,9,10) ,label=c("August", "September", "October"))+
  
  scale_color_discrete(labels = c("ELISA", "Sum of \nMC Congeners")) +
  labs(title = "Total MC from SPATTS") +
  theme_classic()  
```


\newpage


## qPCR

  QPCR is a DNA test to rapidly measure the amount of total cyanobacteria (16s rRNA ) and  toxin genes (mcyE) present. PhytoxigeneTM CyanoDtec test was performed with Applied Biosystem StepOnePlus PCR. Total Cyanobacteria 16s rRNA and toxin gene assay were analyzed in parallel for each month of grab samples. Data for each month is listed in table  below. CyrA and SxtA were not detected  for this year.  The calculated values are expressed as "GeneCopies/mL". 
  
  16s rRNA is a ubiquitis gene which is found in mostly all cyanobacteria. The 16s rRNA gene copies is used to measure relativly how much cyanobacteria is found in your lake. McyE gene is one of the few genes responsible of producing microcystin. Presence of these genes do not indicate that the toxins are present. Detection of McyE gene would indicate that the lake has the potential to produce microcystin and is advised to continue monitoring the lake. We do not become concerned until total cyanobacteria (16s rRNA) are above 200,000 and toxin genes (McyE) are above 10,000.
  

```{r, results='asis', echo=FALSE, fig.align='center'}
third <- MonthOrdered %>% filter(LK_CODE == params$lake, year == params$year) %>%
  select(Mo, '16SRNA', MCYE)

kable(third, format = "markdown", digits = 0, col.names = c("Month", "16s rRNA (copies/mL)", "mcyE (copies/mL)"))



#stargazer(as.data.frame(third),  header=FALSE, covariate.labels = c("Nodularin", "{[D-Asp3]}MC-RR", "MC-RR", "MC-YR", "MC-HtyR", "MC-LR", "{[D-Asp3]}MC-LR", "MC-HilR", "MC-WR", "MC-LA", "MC-LY","MC-LW", "MC-LF", "MC Sum from LC MS/MS ", "MC from ELISA"), notes = "Values are expressed as(mg of MC*${mL^{-1}}$)", notes.align = "r")

```

\newpage

## Nutrient Summary


The chart and table show thye nutrient levels below. The key nutrients we measure are the forms of nitrogen and phosphorous. Phosphorus has long been known to increase the algal and cyanobacteria growth.

```{r  results='asis'}
 aa <- MonthOrdered %>% select(LK_CODE, year, OP:TKN) %>% filter(LK_CODE == params$lake) %>% filter(year == params$year) %>% gather(Nutrients, Level, OP:TKN)

aa %>% ggplot(aes(x=Nutrients, y=Level)) + geom_boxplot() + ylab("Levels in mg/L") + scale_x_discrete(labels=c("Ammonia as N", "Nitrate+Nitrite as N", "Orthophosphate as P", "Total Kjeldahl Nitrogen as N", "Total Phosphorus as P")) + theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r nutrients, results='asis'}
MonthOrdered %>% select(LK_CODE, year, Mo, OP:TKN) %>% filter(LK_CODE == params$lake) %>% select(Mo, OP:TKN) %>% kable(col.names = c("Month", "Orthophosphate (mg P/L)", "Nitrate + Nitrite (mg N/L)", "Ammonia (mg N/L)", "Total Phosphorus (mg P/L)", "Total Kjeldahl Nitrogen (mg N/L)"))
```

## Water Parameters


```{r}
 MonthOrdered %>% filter(LK_CODE == params$lake, year == params$year) %>% 
  select(Mo, wtemp, turb, pH, do, conduc, chloro, phyco) %>%
  #ggplot(aes(x=Mo, y=turb)) + geom_point()
  kable(col.names = c("Month", "Temperature (°C)", "Turbidity (NTU)", "pH", "Dissolve Oxygen (mg/L)", "Conductivity (μS)" , "Chlorophyll-a (RFU)", "Phycocyanin (RFU)"))
```



