---
title: "2017 HABs Data Analysis"
bibliography: sources.bib
csl: citationstyle.csl
output:
  pdf_document:
    toc: true
    number_section: true
    highlight: tango
  #bibliography: source.bib.bib
   # citation_package: biblatex
    
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)


#install.packages("RPostgreSQL")
library(RPostgreSQL)

#Librarys
#install.packages("tidyverse")
library(tidyverse)
#install.packages("shiny")
library(shiny)
#install.packages("GGally")
library(GGally)
#install.packages("viridis")
library(viridis)
#install.packages("ggthemes")
library(ggthemes)
#install.packages("xtable")
library(xtable)

library(readxl)
#install.packages("RColorBrewer")
library(RColorBrewer)
#install.packages("stargazer")
library(stargazer)
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("tensorflow")
#library(tensorflow)
library(rpart)
library(rpart.plot)
# Regression Packages
#install.packages("MASS")
#install.packages("lme4")
#install.packages("leaps")
library(lme4)
library(leaps)
#install.packages("rmarkdown")
library(rmarkdown)
#install.packages("knitr")
library(knitr)
library(kableExtra)
library(leaps)
#install.packages("cowplot")
library(cowplot)
#install.packages("FactoMineR")
#install.packages("factoextra")
#library(MASS)
library(FactoMineR)
library(factoextra)
library(lubridate)

#install.packages("missMDA")
require(missMDA)
#install.packages("texreg")
library(texreg)

#Functions


```

```{r Import Data and some Manipulated tables, include=FALSE}



#Get congeners as zeros and ones by reportable detection limit
RDL_Congeners <- MonthOrdered %>% 
  select(1,2,19:33) %>% 
  mutate(Nodul = if_else(Nodul > 0.03, 1,0)) %>% 
  mutate(D_Asp3_RR = if_else(D_Asp3_RR > 0.03, 1,0)) %>%
  mutate(MC_RR = if_else(MC_RR > 0.03, 1,0)) %>%
  mutate(MC_YR = if_else(MC_YR > 0.03, 1,0)) %>%
  mutate(MC_HTyR = if_else(MC_HTyR > 0.03, 1,0)) %>%
  mutate(MC_LR = if_else(MC_LR > 0.03, 1,0)) %>%
  mutate(D_Asp3_LR = if_else(D_Asp3_LR > 0.03, 1,0)) %>%
  mutate(MC_HiLR = if_else(MC_HiLR > 0.03, 1,0)) %>%
  mutate(MC_WR = if_else(MC_WR > 0.03, 1,0)) %>%
  mutate(MC_LA = if_else(MC_LA > 0.03, 1,0)) %>%
  mutate(MC_LY = if_else(MC_LY > 0.03, 1,0)) %>%
  mutate(MC_LW = if_else(MC_LW > 0.03, 1,0)) %>%
  mutate(MC_LF = if_else(MC_LF > 0.03, 1,0)) %>%
  mutate(SUM = if_else(SUM > 0.03, 1,0)) %>%
  mutate(ELISA = ifelse(ELISA > 0.15, 1,  0))
  #This is an example to keep all other values the same
  #mutate(ELISA = case_when(ELISA > 0.5 ~ 1, TRUE ~ ELISA))

# Subset for PCA, no log transformation and only all numerical values
MODEL_PCA <- MonthOrdered %>%
  select(8:12, 14, 15, 16, 32, 52, 55:61, 7, 4)

#[MODEL_PCA2 <-  MonthOrdered  %>% select(8:12, 14, 15, 16, 32, 52, 55:61, 7, 4, 65)

# Subset with NA removed
MODEL_NA <- MODEL %>% select(1:6, SUM, 37:43, 44) %>% na.omit()

# Logistic Model

LogModel <- MODEL %>% mutate(MS_SUM = ifelse(SUM > 0.1, 1, 0))



# LogTransformed Averages for each lake code         
lkcode_grp <- MonthOrdered %>%
  group_by(LK_CODE) %>% select(LK_CODE, OP:phyco) %>%
  summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  mutate_at(vars(OP:TKN, Nodul:S_ELISA), funs(log1p(.))) %>%
  mutate_at(vars("X16SRNA", MCYE, do:phyco), funs(log(.)))



# Subsets for scatterplot matrix. Log transformed
grabNutes <- MonthOrdered %>% 
  select(LK_CODE, month, NOX, OP, NH3, TP, TKN, ELISA, SUM,TNTP) %>% 
  mutate_at(vars(NOX, OP, NH3, TP, TKN, ELISA, SUM), funs(log1p(.)))

spattNutes <- MonthOrdered %>% 
  select(LK_CODE, month, NOX, OP, NH3, TP, TKN, S_ELISA, S_SUM) %>% 
  mutate_at(vars(NOX, OP, NH3, TP, TKN, S_ELISA, S_SUM), funs(log1p(.)))

cong_lk <- MonthOrdered %>%
  select(LK_CODE, Nodul:ELISA) %>% 
  group_by(LK_CODE) %>% summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  rowwise() %>% 
  mutate(sumz = sum(Nodul,D_Asp3_RR, MC_RR, MC_YR, MC_HTyR, MC_LR, D_Asp3_LR, MC_HiLR, MC_WR, MC_LA, MC_LY, MC_LW, MC_LF, na.rm = TRUE)) %>% 
  mutate_at(vars(Nodul:MC_LF), funs(./sumz)) %>% 
  mutate_at(vars(Nodul:MC_LF), funs(.*100))



# Tables for congeners barplots as percentages


landus_sum <- LOGTransformed %>% gather(LandUse, percentage, Water:Wetlands)





cong_2 <- inner_join(
  cong_sum_sp, Lake_info_comprehensive, by = "LK_CODE"
  ) %>% 
  select(LK_CODE, congener, percentage, Lat)
#
```







*Coordinates are the location of sampling point in NAD83

\newpage

# Distribution/Visual Explore

## Nutrients

```{r, fig.height=4, fig.width=10, fig.align='left'}
first.half <- UltimateFactor %>% select(OP:TKN, TN)

PLOTNUT1 <- first.half %>% 
  gather(Nutrients, count, OP:NH3) %>% 
  ggplot(aes((count), fill=Nutrients)) + 
  geom_density(alpha=0.8)  + 
  scale_fill_viridis( discrete = TRUE,labels = c("Ammonia", "Nitrate+Nitrite", "Orthophosphate")) + 
  scale_x_continuous(limits = (c(0,0.25))) + 
  theme_classic() + 
  ylab("count") + 
  xlab("ppm or mg X/L") +
  geom_vline(xintercept=mean(first.half$NH3, na.rm = TRUE)) +
  geom_vline(xintercept=median(first.half$OP, na.rm = TRUE)) +
  geom_vline(xintercept=median(first.half$NOX, na.rm = TRUE))

PLOTNUT2 <- first.half %>% gather(Nutrients, count, TP:TKN) %>% 
  ggplot(aes((count), fill=Nutrients)) + 
  geom_density(alpha=0.8)  + 
  scale_fill_viridis(discrete = TRUE, option = "B", labels = c("Total Kjeldahl Nitrogen", "Total Phosphorus")) + 
  scale_x_continuous(limits = (c(0,0.80))) + 
  scale_y_continuous(limits = (c(0,75))) + 
  theme_classic() + 
  ylab("count") + 
  xlab("ppm or mg X/L") +
  geom_vline(xintercept=mean(first.half$TP, na.rm = TRUE)) +
  geom_vline(xintercept=mean(first.half$TKN, na.rm = TRUE))

plot_grid(PLOTNUT1, PLOTNUT2)
```


```{r, qpcr plot,}
MonthOrdered %>% gather(Gene, Copies, 'X16SRNA', MCYE) %>%
  ggplot(aes(x=Gene, y=log(Copies))) + geom_boxplot() + theme_classic()

```

\newpage

## Mass Spec and ELISA






 



## Congeners from Grab Samples

```{r 11, warning=FALSE, echo = FALSE}
cong_sum %>% ggplot(aes(x=LK_CODE, y=percentage, fill=congener)) + geom_bar(stat="identity") + scale_fill_viridis(discrete=TRUE, option="B"
                                                                                                                  ,label=c( "[D-Asp3] MC-LR", "[D-Asp3] MC-RR",   "MC-HilR", "MC-HtyR","MC-LA", "MC-LF", "MC-LR", "MC-LW", "MC-LY","MC-RR", "MC-WR", "MC-YR", "Nodularin")
                                                                                                                  ) + theme_classic() + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.4)) + xlab("Lake") +
  #labs(title="Proportion of Congeners in Grab Sample", fill="Congener") +
  ylab("%")

```




## Congeners by Latitude






## SPATTS Congeners by Lake


## SPATTS Congeners by Latitude

```{r 31, warning=FALSE, echo = FALSE}
cong_2 %>% ggplot(aes(x=Lat, y=percentage, fill=congener, width=0.09)) + geom_bar(stat="identity")  + scale_fill_viridis(discrete=TRUE, option="B"
                                                                                                                         ,label=c( "[D-Asp3] MC-LR", "[D-Asp3] MC-RR",   "MC-HilR", "MC-HtyR","MC-LA", "MC-LF", "MC-LR", "MC-LW", "MC-LY","MC-RR", "MC-WR", "MC-YR", "Nodularin")
                                                                                                                         ) + 
  #scale_y_continuous(limits = c(0,100)) + 
  coord_flip() + theme_classic()

```




\newpage

## Land Use by Lake

```{r 34, warning=FALSE, echo = FALSE}
landus_sum %>% ggplot(aes(x=LK_CODE, y=percentage, fill=LandUse)) + theme_classic() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))  + geom_bar(stat="identity") + scale_fill_manual(values=c("#ffff00",	"#804d33",	"#ff0000",	"#006600",	"#e5cc99",	"#b2b200",	"#0000ff",	"#00ffff"
)) + xlab("Lake") + 
  #labs(title="Land Use % by Lake") + 
  ylab("%")

```



# Matrix-Exploratory analysis


## Nutrients

```{r scatter2, message=FALSE, warning = FALSE, echo = FALSE, fig.width=11, fig.height=10}
ggpairs(MODEL, columns=c( "NOX","NH3", "OP", "TKN", "TP", "SUM"), columnLabels = c("Nitrate+Nitrite", "Ammonia", "Orthophosphate", "Total Nitrogen", "Total Phosphorus", "MC Sum from LC MS/MS"), upper = list(continuous = wrap("cor", na.rm = TRUE, use = "complete.obs", size = 2.5))) + theme_classic()
```


## Nutrients-By each month

```{r cars, message=FALSE, warning = FALSE, echo = FALSE, fig.width=11, fig.height=10}
ggpairs(grabNutes,
        aes(colour=month), 
        columns=c("NOX","NH3", "OP", "TKN", "TP", "month","SUM"), columnLabels = c("Nitrate+Nitrite", "Ammonia", "Orthophosphate", "Total K Nitrogen", "Total Phosphorus", "Month","MC Sum from LC MS/MS"), 
        upper = list(continuous = wrap("cor", size = 2.5))) + theme_classic()
```


## SPATTS

```{r scatter3, message=FALSE, warning = FALSE, echo = FALSE, fig.width=10, fig.height=10}
ggpairs(MODEL, columns=c("NOX", "NH3", "OP", "TP", "TKN", "S_SUM"), columnLabels = c("Nitrate+Nitrite", "Ammonia", "Orthophosphate", "Total Phosphate", "Total Nitrogen", "Sum of MC from MS/MS"), upper = list(continuous = wrap("cor", size = 2.5))) + theme_classic()

```

## Land Use 

Regression with land use and morphology, 


## Land Use correlation with all lakes

```{r test12, message=FALSE, warning = FALSE, echo = FALSE, fig.width=10, fig.height=10}
LandUseGGPAIRS1 <- model1 %>% select(Water:Wetlands, SUM) 
ggpairs(LandUseGGPAIRS1, columnLabels = c("Water", "Developed", "Barren", "Forest", "Shrubs", "Herbaceous", "Agriculture", "Wetlands", "Sum of MC ")) + theme_classic()
```

`r nrow(LandUseGGPAIRS1)` of observations

## Land Use Correlation with open lakes

```{r test, message=FALSE, warning = FALSE, echo = FALSE, fig.width=10, fig.height=10}
 model1 %>% filter(Network > 0) %>% filter(Network %in% c(3,4)) %>% select(Water:Wetlands, SUM) %>% ggpairs() + theme_classic()
```






# R Code 

## R Setup

**#Use RStudio or install R packages.** 
**#Input these commands into console to install the suggested packages.**


```
#Library and packages setup
library(ggplot2)
library("broom")
library(tidyverse)
library(knitr)
library(dplyr)
library(GGally)
library(viridis)
library(readxl)
library(psych)
library(xtable)
library(rmarkdown)
library(readxl)
library(tidyr)
library(RColorBrewer)
```



**Import data from excel files. Get the mastersheet excel file from gdrive. Import it either graphically from file, import datashets, excele and put the option of -999 as na value. Otherwise use these commands:**

```
#Load Data--change where your path to file is!
Master_Sheet <- read_excel("Path_to_file.xlsx", 
                                     na = "-999")
Lake_info_comprehensive <- read_excel("Path_to_file.xlsx")
```

**#Manipulate and create subset data. This uses the pipe function %>% which is like saying "and then". Reordered the data by month and sorted as levels. Then various subset data is generated. Log (1+x) for x values generally that are low. log1p is an r function for that transformation:**

```
MonthOrdered <- Master_Sheet %>% 
  mutate(month = factor(month.name[Month], levels = month.name)) %>% 
  arrange(month)

info <- Lake_info_comprehensive %>% 
  rowwise() %>% 
  mutate(sum = sum(Water,Developed,Barren,Forest,Shrubs,
                   Herbaceous,Agriculture,Wetlands)) %>% 
  mutate_at(vars(Water:Wetlands), funs(./sum)) %>% 
  mutate_at(vars(Water:Wetlands), funs(.*100))
         
lkcode_grp <- MonthOrdered %>%
  group_by(LK_CODE) %>% select(LK_CODE, OP:phyco) %>%
  summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  #mutate_at(vars(OP:TKN, Nodul:S_ELISA), funs(log1p(.))) %>%
  mutate_at(vars("16SRNA", MCYE, do:phyco), funs(log(.)))

model <- full_join(info, lkcode_grp, by = "LK_CODE")

model1 <- full_join(info, lkcode_grp, by = "LK_CODE") %>% 
  mutate_at(vars(Water:Wetlands), funs(./100))

fullmod <- full_join(info, MonthOrdered, by ="LK_CODE") %>%
  mutate_at(vars(Water:Wetlands), funs(./100))

grabNutes <- MonthOrdered %>% 
  select(LK_CODE, month, NOX, OP, NH3, TP, TKN, ELISA, SUM,TNTP) %>% 
  mutate_at(vars(NOX, OP, NH3, TP, TKN, ELISA, SUM), funs(log1p(.)))

spattNutes <- MonthOrdered %>% 
  select(LK_CODE, month, NOX, OP, NH3, TP, TKN, S_ELISA, S_SUM) %>% 
  mutate_at(vars(NOX, OP, NH3, TP, TKN, S_ELISA, S_SUM), funs(log1p(.)))

cong_lk <- MonthOrdered %>%
  select(LK_CODE, Nodul:ELISA) %>% 
  group_by(LK_CODE) %>% summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  rowwise() %>% 
  mutate(sumz = sum(Nodul,D_Asp3_RR, MC_RR, MC_YR, MC_HTyR, MC_LR, D_Asp3_LR, MC_HiLR, MC_WR, MC_LA, MC_LY, MC_LW, MC_LF, na.rm = TRUE)) %>% 
  mutate_at(vars(Nodul:MC_LF), funs(./sumz)) %>% 
  mutate_at(vars(Nodul:MC_LF), funs(.*100))

cong_sp <- MonthOrdered %>%
  select(LK_CODE, S_D_Asp3_RR:S_ELISA) %>% 
  group_by(LK_CODE) %>% summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  rowwise() %>% 
  mutate(sumz = sum(S_Nodul,S_D_Asp3_RR, S_MC_RR, S_MC_YR, S_MC_HtyR, S_MC_LR, S_D_Asp3LR, S_MC_HilR, S_MC_WR, S_MC_LA, S_MC_LY, S_MC_LW, S_MC_LF, na.rm = TRUE)) %>% 
  mutate_at(vars(S_D_Asp3_RR:S_MC_LF), funs(./sumz)) %>% 
  mutate_at(vars(S_D_Asp3_RR:S_MC_LF), funs(.*100))

cong_sum <- cong_lk %>% gather(congener, percentage, Nodul:MC_LF)

landus_sum <- model %>% gather(LandUse, percentage, Water:Wetlands)

cong_sum_sp <- cong_sp %>% gather(congener, percentage, S_D_Asp3_RR:S_MC_LF)

cong_1 <- inner_join(
  cong_sum, Lake_info_comprehensive, by = "LK_CODE"
  ) %>% 
  select(LK_CODE, congener, percentage, Lat)

cong_2 <- inner_join(
  cong_sum_sp, Lake_info_comprehensive, by = "LK_CODE"
  ) %>% 
  select(LK_CODE, congener, percentage, Lat)
```
## R Graphs

**Make awesome plots with ggplot, ggpairs, or other extraneous things here. Each block generates unique plots:**

```
ggpairs(grabNutes, aes(color=month), columns=c("NOX", "month", "OP", "TKN", "TP", "SUM"), upper = list(continuous = wrap("cor", size = 2.5)))

cong_1 %>% ggplot(aes(x=Lat, y=percentage, fill=congener, width=0.09)) + geom_bar(stat="identity")  + scale_fill_viridis(discrete=TRUE, option="magma") + scale_y_continuous(limits = c(0,100)) + coord_flip() + theme_classic()

fit1 <- lm(SUM ~  Developed, data=fullmod)
summary(fit1)

```


# References
