---
title: "HABS 2017-Data Analysis"

output:
  beamer_presentation: default
  ioslides_presentation: default
  slidy_presentation: default
csl: citationstyle.csl
bibliography: sources.bib
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

#Librarys
#install.packages("tidyverse")
library(tidyverse)
#install.packages("shiny")
library(shiny)
#install.packages("GGally")
library(GGally)
#install.packages("viridis")
library(viridis)
#install.packages("ggthemes")
library(ggthemes)
#install.packages("xtable")
library(xtable)

library(readxl)
#install.packages("RColorBrewer")
library(RColorBrewer)
#install.packages("stargazer")
library(stargazer)
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("tensorflow")
#library(tensorflow)
library(rpart)
library(rpart.plot)
# Regression Packages
#install.packages("MASS")
#install.packages("lme4")
#install.packages("leaps")
library(lme4)
library(leaps)
#install.packages("rmarkdown")
library(rmarkdown)
#install.packages("knitr")
library(knitr)
library(kableExtra)
library(leaps)
#install.packages("cowplot")
library(cowplot)
#install.packages("FactoMineR")
#install.packages("factoextra")
#library(MASS)
library(FactoMineR)
library(factoextra)

#Functions







#Load Data

Master_Sheet <- read_excel("20170214_Master_Sheet.xlsx", 
                                     na = "-999")
Lake_info_comprehensive <- read_excel("C:/Users/John Conner/Google Drive/2017-2018 HAB Survey/DATA/Master/Lake_info_comprehensive.xls")

#Manipulation------

# Order months by month instead of alphabetical
MonthOrdered <- Master_Sheet %>% 
  mutate(month = factor(month.name[Month], levels = month.name)) %>% 
  arrange(month)

# Subset for generalized model and apply log transformation
MODEL <- MonthOrdered %>% 
  select(8:12, 14, 16, 20:33, 39:53, 55:61) %>%
  mutate_at(vars(OP:TKN, 8:36), funs(log1p(.)))

# Subset for PCA, no log transformation and only all numerical values
MODEL_PCA <- MonthOrdered %>%
  select(8:12, 14, 15, 16, 32, 52, 55:61)

# Subset with NA removed
MODEL_NA <- MODEL %>% select(1:6, SUM, 37:43) %>% na.omit()

# Calculate land use percentages
info <- Lake_info_comprehensive %>% 
  rowwise() %>% 
  mutate(sum = sum(Water,Developed,Barren,Forest,Shrubs,
                   Herbaceous,Agriculture,Wetlands)) %>% 
  mutate_at(vars(Water:Wetlands), funs(./sum)) %>% 
  mutate_at(vars(Water:Wetlands), funs(.*100))

# Averages for each lake code         
lkcode_grp <- MonthOrdered %>%
  group_by(LK_CODE) %>% select(LK_CODE, OP:phyco) %>%
  summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  mutate_at(vars(OP:TKN, Nodul:S_ELISA), funs(log1p(.))) %>%
  mutate_at(vars("16SRNA", MCYE, do:phyco), funs(log(.)))

# Joined 
model <- full_join(info, lkcode_grp, by = "LK_CODE")

model1 <- full_join(info, lkcode_grp, by = "LK_CODE") %>% 
  mutate_at(vars(Water:Wetlands), funs(./100))

fullmod <- full_join(info, MonthOrdered, by ="LK_CODE") %>%
  mutate_at(vars(Water:Wetlands), funs(./100))

# Subsets
grabNutes <- MonthOrdered %>% 
  select(LK_CODE, month, NOX, OP, NH3, TP, TKN, ELISA, SUM,TNTP) %>% 
  mutate_at(vars(NOX, OP, NH3, TP, TKN, ELISA, SUM), funs(log1p(.)))

spattNutes <- MonthOrdered %>% 
  select(LK_CODE, month, NOX, OP, NH3, TP, TKN, S_ELISA, S_SUM) %>% 
  mutate_at(vars(NOX, OP, NH3, TP, TKN, S_ELISA, S_SUM), funs(log1p(.)))

cong_lk <- MonthOrdered %>%
  select(LK_CODE, Nodul:ELISA) %>% 
  group_by(LK_CODE) %>% summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  rowwise() %>% 
  mutate(sumz = sum(Nodul,D_Asp3_RR, MC_RR, MC_YR, MC_HTyR, MC_LR, D_Asp3_LR, MC_HiLR, MC_WR, MC_LA, MC_LY, MC_LW, MC_LF, na.rm = TRUE)) %>% 
  mutate_at(vars(Nodul:MC_LF), funs(./sumz)) %>% 
  mutate_at(vars(Nodul:MC_LF), funs(.*100))

cong_sp <- MonthOrdered %>%
  select(LK_CODE, S_D_Asp3_RR:S_ELISA) %>% 
  group_by(LK_CODE) %>% summarise_all(funs(mean(.,na.rm = TRUE))) %>%
  rowwise() %>% 
  mutate(sumz = sum(S_Nodul,S_D_Asp3_RR, S_MC_RR, S_MC_YR, S_MC_HtyR, S_MC_LR, S_D_Asp3LR, S_MC_HilR, S_MC_WR, S_MC_LA, S_MC_LY, S_MC_LW, S_MC_LF, na.rm = TRUE)) %>% 
  mutate_at(vars(S_D_Asp3_RR:S_MC_LF), funs(./sumz)) %>% 
  mutate_at(vars(S_D_Asp3_RR:S_MC_LF), funs(.*100))

# Tables for barplots as percentages
cong_sum <- cong_lk %>% gather(congener, percentage, Nodul:MC_LF)

landus_sum <- model %>% gather(LandUse, percentage, Water:Wetlands)

cong_sum_sp <- cong_sp %>% gather(congener, percentage, S_D_Asp3_RR:S_MC_LF)

cong_1 <- inner_join(
  cong_sum, Lake_info_comprehensive, by = "LK_CODE"
  ) %>% 
  select(LK_CODE, congener, percentage, Lat)

cong_2 <- inner_join(
  cong_sum_sp, Lake_info_comprehensive, by = "LK_CODE"
  ) %>% 
  select(LK_CODE, congener, percentage, Lat)
#

```

```{r new, include=FALSE}
library(lavaan)
library(semPlot)
PCA_Land <- model1 %>% select(8, 11:21, 23:29,  47, 67) %>% na.omit()
```

## Outline

- Basic stats
- Lets Explore into the wilderness
- Modeling
    a. Multiple Linear Regression
    b. Principal Component Analysis
    c. Regression Classification Tree
    d. Structural Equation Modeling
 
    


## Modeling Assumptions

- We are assuming linear relationships between dependent and independent variables
- The samples taken are enough to represent this relationship

## Errors and Fallacies

- Covariants that can be closely related
- Sampling size and selection
- Regional effects



## Explore

- Maybe we can find something else in our data
- Lets explore



## Nutrients

```{r, results='asis', echo=FALSE}
first.half <- MonthOrdered %>% select(OP:TKN)
stargazer(as.data.frame(first.half), header=FALSE, covariate.labels = c(" Orthophosphate (mg P/mL)","Nitrate+Nitrite (mg N/mL)", "Ammonia (mg N/mL) ", "Total Phosphorus (mg P/mL)", "Total Nitrogen (mg N/mL)"), font.size = "small")

```



## Nutrients


```{r, fig.height=7}
first.half %>% gather(Nutrients, count, OP:NH3) %>% ggplot(aes((count), fill=Nutrients)) + geom_density(alpha=0.8)  + scale_fill_brewer(palette = "Dark2" , labels = c("Ammonia", "Nitrate+Nitrite", "Orthophosphate")) + scale_x_continuous(limits = (c(0,0.25))) + theme_classic()
```

## Nutrients

```{r, fig.height=7}
first.half %>% gather(Nutrients, count, TP:TKN) %>% ggplot(aes((count), fill=Nutrients)) + geom_density(alpha=0.8)  + scale_fill_brewer(palette = "Set3" , labels = c("Total Phosphate", "Total Nitrogen")) + scale_x_continuous(limits = (c(0,0.25))) + scale_y_continuous(limits = (c(0,75))) + theme_classic()

```


## QPCR-Summary


```{r,warning = FALSE, echo=FALSE, message=FALSE, header=FALSE, results='asis'}
second <- MonthOrdered %>%
  select("16SRNA", MCYE:SxtA)
stargazer(as.data.frame(second), header = FALSE,  digits = 0, covariate.labels = c("16S rRNA", "mcyE", "cyrA", "sxtA"))

```


## QPCR spread

```{r, fig.height=7}
MonthOrdered %>% gather(Gene, Copies, '16SRNA', MCYE) %>% 
ggplot(aes(((Copies)), fill=Gene)) + geom_density(alpha=0.8)  + scale_fill_brewer(palette = "Dark2" ) + theme_classic()
```


## QPCR

```{r, fig.height=7}
MonthOrdered %>% gather(Gene, Copies, '16SRNA', MCYE) %>% 
ggplot(aes((log(Copies)), fill=Gene)) + geom_density(alpha=0.8)  + scale_fill_brewer(palette = "Dark2" ) + theme_classic()

```

## Mass Spec- Grab Samples



```{r boxlake, warning=FALSE, echo=FALSE, fig.height=7 }
MonthOrdered %>% ggplot (aes(x=LK_CODE, y=(SUM))) + geom_boxplot() + theme_classic() + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.4))

```
 
 
## ELISA-Grab sample


```{r boxlake2, warning=FALSE, echo=FALSE, fig.height=7}
MonthOrdered %>% ggplot (aes(x=LK_CODE, y=(ELISA))) + geom_boxplot() + theme_classic() + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.4))

```


## SPATTS

```{r boxlake33, warning=FALSE, echo=FALSE, fig.height=7 }

MonthOrdered %>% ggplot (aes(x=LK_CODE, y=((S_SUM)))) + geom_boxplot() + theme_classic() + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.4))

```


## Congener Summary

```{r 11, warning=FALSE, echo = FALSE}
cong_sum %>% ggplot(aes(x=LK_CODE, y=percentage, fill=congener)) + geom_bar(stat="identity") + scale_fill_viridis(discrete=TRUE, option="magma") + theme_classic() + theme(axis.text.x = element_text(angle = 90, size = 6, vjust = 0.4)) + xlab("Lake") + labs(title="Proportion of Congeners in each Lake", fill="Congener") + ylab("%")

```

## Congener

```{r, results='asis', echo=FALSE}
third <- MonthOrdered %>%
  select(Nodul:ELISA)
stargazer(as.data.frame(third), header=FALSE, covariate.labels = c("Nodularin", "{[D-Asp3]}MC-RR", "MC-RR", "MC-YR", "MC-HtyR", "MC-LR", "{[D-Asp3]}MC-LR", "MC-HilR", "MC-WR", "MC-LA", "MC-LY","MC-LW", "MC-LF", "MC Sum from LC MS/MS ", "MC from ELISA"), notes = "Values are expressed as (mg of MC/mL)", notes.align = "r", font.size = "tiny")

```


## Congener

```{r 2, warning=FALSE, echo = FALSE}
cong_sum_sp %>% ggplot(aes(x=LK_CODE, y=percentage, fill=congener)) + theme_classic() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))  + geom_bar(stat="identity") + scale_fill_viridis(discrete=TRUE, option="magma") + xlab("Lake") + labs(title="Congeners % by Lake from SPATTS", fill="Congener") + ylab("%")

```
 
## Congener-SPATTS


```{r, results='asis', warning = FALSE, echo=FALSE, message=FALSE}
fourth <- MonthOrdered %>%
  select(39:53)
stargazer(as.data.frame(fourth), header=FALSE, covariate.labels = c("{[D-Asp3]}MC-RR", "MC-RR", "Nodularin", "MC-YR", "MC-HtyR", "MC-LR", "{[D-Asp3]}MC-LR", "MC-HilR", "MC-WR", "MC-LA", "MC-LY","MC-LW", "MC-LF", "MC Sum from LC MS/MS", "MC from ELISA"),  notes = "Values are expressed as (mg of MC*${mL^{-1}}$)", notes.align = "r", font.size = "tiny")

```

 
## Matrix-Nutrients

```{r scatter2, message=FALSE, warning = FALSE, echo = FALSE, fig.width=13, fig.height=10}
ggpairs(MODEL, columns=c( "NOX","NH3", "OP", "TKN", "TP", "SUM"), columnLabels = c("Nitrate+Nitrite", "Ammonia", "Orthophosphate", "Total Nitrogen", "Total Phosphorus", "MC Sum from LC MS/MS"), upper = list(continuous = wrap("cor", na.rm = TRUE, use = "complete.obs", size = 7))) + theme_classic()
```

## Matrix-Land Use-Grab

```{r test12, message=FALSE, warning = FALSE, echo = FALSE, fig.width=13, fig.height=10}
LandUseGGPAIRS1 <- model1 %>% select(Water:Wetlands, SUM) 
ggpairs(LandUseGGPAIRS1,  upper = list(continuous = wrap("cor", size = 7))) + theme_classic()
```

## Matrix-SPATTS

```{r scatter3, message=FALSE, warning = FALSE, echo = FALSE, fig.width=13, fig.height=10}
ggpairs(MODEL, columns=c("NOX", "NH3", "OP", "TP", "TKN", "S_SUM"), columnLabels = c("Nitrate+Nitrite", "Ammonia", "Orthophosphate", "Total Phosphate", "Total Nitrogen", "Sum of MC from MS/MS"), upper = list(continuous = wrap("cor", size = 7))) + theme_classic()

```

## Matrix-LandUse on SPATTS

```{r scatter55, message=FALSE, warning = FALSE, echo = FALSE, fig.width=13, fig.height=10}
LandUseGGPAIRS1 <- model1 %>% select(Water:Wetlands, S_SUM) 
ggpairs(LandUseGGPAIRS1, upper = list(continuous = wrap("cor", size = 7))) + theme_classic()

```




## Stepwise Regression

Start:  AIC=-187.16

SUM ~ OP + NOX + NH3 + TP + TKN + `16SRNA` + wtemp + pH + conduc + 
    turb + chloro + phyco

Step:  AIC=-189.1

SUM ~ OP + NOX + NH3 + TP + `16SRNA` + wtemp + pH + conduc + 
    turb + chloro + phyco
    
## Comparision of linear fits

```{r, results='asis'}
fit1 <- glm(SUM ~ OP + NOX + NH3 + TP + TKN  + wtemp + pH + conduc + 
    turb + chloro + phyco, family = gaussian(), MODEL_NA)



fit3 <- glm(SUM ~ OP  + pH + turb + chloro + phyco, family = gaussian(), MODEL)

fit4 <- glm(SUM ~ OP  + pH + turb + phyco,  family = gaussian(), MODEL)

fit5 <- glm(SUM ~ OP + '16SRNA' + turb, family = gaussian(), MODEL)

stargazer(fit1, fit3, fit4, fit5, header=FALSE, single.row = TRUE, column.sep.width = "1pt", font.size = "tiny")
```   

## Land use


```{r, results='asis'}
fit6 <- glm(SUM ~ Wetlands + Herbaceous + Developed + Water + Shrubs + Barren, family = gaussian(), PCA_Land)



fit7 <- glm(SUM ~ Developed + Agriculture + Wetlands,   family = gaussian(), PCA_Land)

fit8 <- glm(SUM ~  Developed + Wetlands + Shrubs + Water, family = gaussian(), PCA_Land)

fit9 <- glm(SUM ~ Developed + Wetlands + Shrubs , family = gaussian(), PCA_Land)

stargazer(fit6, fit7, fit8, fit9, header=FALSE, single.row = TRUE, column.sep.width = "1pt", font.size = "tiny")
```   


## Regression

```{r, results='asis'}
fit6 <- glm(S_SUM ~ OP + NOX + NH3 + TP + TKN , family = gaussian(), PCA_Land)



fit7 <- glm(S_SUM ~ OP + Developed + Agriculture + Wetlands,   family = gaussian(), PCA_Land)

fit8 <- glm(S_SUM ~ OP + Developed + Wetlands,   family = gaussian(), PCA_Land)

fit9 <- glm(S_SUM ~ Developed + Wetlands + Agriculture,  family = gaussian(), PCA_Land)

stargazer(fit6, fit7, fit8, fit9, header=FALSE, single.row = TRUE, column.sep.width = "1pt", font.size = "tiny")
```   

## PCA

### Purpose 

- Explore the data
- Reduce the dimensionality of data
- Redeuce noise and redundancy
- Identify correlated variables

## Preparing Data

- Performing principal component analysis (PCA) must be done on a normalized dataset. 
- The data is scaled and centered to the variance such that the axis of each variable are not so different. 
- Data is standardized by this formula:

$$\frac{x_{i} - mean(x)}{\sigma}$$






```{r}

PCA_data2 <- PCA(PCA_Land, scale = TRUE, graph =  FALSE)
#PCA_data_center <- PCA(MODEL, ncp = ncol (MODEL), scale.unit="FALSE")

```












## PCA-Grab

```{r}
fviz_pca_var(PCA_data,
             
             repel = TRUE # Avoid text overlapping
             )
```


## PCA-Land Use

```{r}
fviz_pca_var(PCA_data2,
             
             repel = TRUE # Avoid text overlapping
             )

```

## Regression Trees-Grab sample


```{r}


rpart1 <- rpart(SUM ~ TKN + TP +  NOX + NH3 + doy +  turb, data = MODEL_NA)

rpruned1 <- rpart
rpart.plot(rpart1)


```

## Cost complexity vs tree size

```{r prune}
plotcp(rpart1)
```



## Regression Trees - SPATTS

```{r}


rpart2 <- rpart(S_SUM ~ TKN + TP + phyco + NOX + NH3 + OP +  turb, data = MODEL_PCA, minsplit=3)

rpart.plot(rpart2)


```

## Regression trees- Land Use with grab sample
```{r}


rpart3 <- rpart(SUM ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm, data = PCA_Land, minsplit=4)

rpart.plot(rpart3)


```

## Regression Trees - Land use with SPATTS

```{r}


rpart4 <- rpart(S_SUM ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm, data = PCA_Land, minsplit=4)

rpart.plot(rpart4)


```

## Structural Path Analysis


```{r}
model.nutrients121 <- ' SUM ~ OP + NOX + NH3 + TKN + TP + phyco + turb

'
                   
                    
fit <- sem(model.nutrients121, data=MonthOrdered)

                    
                    
semPaths(fit, what = "stand", style = "lisrel", layout = "spring", bg = "#000000", edge.label.cex = 1.1 )
```

## Path Analysis-Land Use

```{r}
model.nutrients232 <- ' SUM ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm
'
                   
                    
fit <- sem(model.nutrients232, data=PCA_Land)

                    
                    
semPaths(fit, what = "stand", style = "lisrel", layout = "spring", bg = "#000000", edge.label.cex = 1.1 )
```

## SPATT Land Use Path analysis

```{r}

model.nutrients33 <- ' S_SUM ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm
'
                   
                    
fit <- sem(model.nutrients33, data=PCA_Land)

                    
                    
semPaths(fit, what = "stand", style = "lisrel", layout = "spring", bg = "#000000", edge.label.cex = 1.1 )
```

## Path Analysis-Land Use with nutrients

```{r}
model.nutrients441 <- ' OP ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm

NOX ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm

NH3 ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm


'
                   
                    
fit <- sem(model.nutrients441, data=PCA_Land)

                    
                    
semPaths(fit, what = "stand", style = "lisrel", layout = "spring", bg = "#000000", edge.label.cex = 1.1 )
```

## Total P and N Path analysis 

```{r}
model.nutrients4441 <- ' TP ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm

TKN ~ Water + Developed + Shrubs + Barren + Herbaceous + Wetlands + Agriculture + Max_Depth + Lake_Area_sqKm + Watershed_Area_sqKm

'
                   
                    
fit <- sem(model.nutrients4441, data=PCA_Land)

                    
                    
semPaths(fit, what = "stand", style = "lisrel", layout = "spring", bg = "#000000", edge.label.cex = 1.1 )
```
 




